---
title: "Boston Buoy Data Analysis Report"
author: "Danping Liu"
date: "9/24/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(lubridate)
library(dplyr)
library(ggplot2)
library(gridExtra)
```
### Introduction 
The objective for this project is to find the evidence for global warming. From NASA's *Climate change evidence: How do we know?* page, it is said that rising average temperature, warming oceans and extreme climate are significant factors that show global warming. So to prove that global warming is true, we are going to analyze the change in air temperature and water temperature. 

Firstly, we import the data we are going use to R. Secondly, we build linear regression models for to check how the monthly temperature change.Thirdly, we build linear regression models to check how the annual average temperature change. Lastly, we make conclusion based on the two aspects above. 

### Import data
The data we use is NOAA's Boston's air and water temperature from 1987 to 2016. There are 246245 observations in total. For better comparison, we remove the NA values and use the temperature at 12pm everyday. As a result, we have 10146 observations for analyzing. The year, month, day, hour, air temperature and water temperature are stored in the data frame as string. For later analyzing, we change the data types for air and water temperature to double. We create a new TIME column combining the year, month, day and hour columns into POXISct form. However, POXISct data cannot join the regression fitting, so we unclass it to get a double type variable representing the number of seconds from 00:00:00 January 1, 1970. 
```{r}
load("MR.Rdata")
# Remove NA values
`%notin%` <- Negate(`%in%`)
MRsub <- filter(MR, ATMP%notin%c(999,99,"999","99","999.0","99.0") & WTMP%notin%c(999,99,"999","99","999.0","99.0") & hh=="12")
# Convert string values to double
MRsub$ATMP <- as.double(MRsub$ATMP)
MRsub$WTMP <- as.double(MRsub$WTMP)
# Convert YYYY, MM, DD and hh to POSIXct and then to seconds from 1970/1/1
buoyTimeStr <- paste(MRsub$YYYY, MRsub$MM, MRsub$DD, MRsub$hh, sep="-")
MRsub$TIME <- ymd_h(buoyTimeStr)
MRsub$DATE <- unclass(MRsub$TIME)
MRsub$YYYY <- as.double(MRsub$YYYY)
```

### Monthly temperature change
Since the temperature varies during different months, we decide to build linear regression models for each month respectively and for annual average. For better understanding, we make a table presenting how monthly air temperature and water temperature is changing from year to year. We also make plots of temperature versus time for every month. 
```{r}
numMonths <- levels(factor(MRsub$MM))
nameMonths <- c("January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December")
fitA_list <- list()
fitW_list <- list()
coefAPerYear <- list()
coefWPerYear <- list()
for(i in 1:12){
  fitA_list[[i]] <- lm(ATMP~DATE, data=MRsub[MRsub$MM==numMonths[i],])
  fitW_list[[i]] <- lm(WTMP~DATE, data=MRsub[MRsub$MM==numMonths[i],])
  coefAPerYear <- 60*60*24*365*coef(fitA_list[[i]])[2]
  coefWPerYear <- 60*60*24*365*coef(fitW_list[[i]])[2]
}
MRmonth <- data.frame(nameMonths, coefAPerYear, coefWPerYear)
kable(MRmonth, col.names=c("Month", 
                           "Monthly air temperature change every year in celcius degrees", 
                           "Monthly water temperature change every year in celcius degrees"))
plotFit <- ggplot(data=MRsub)+
  geom_point(aes(x=DATE,y=ATMP),alpha=0.1,color="black")+
  geom_point(aes(x=DATE,y=WTMP),alpha=0.1, color = "#56B4E9")+
  geom_smooth(aes(x=DATE,y=ATMP,color="Air"),method="lm", formula="y~x")+
  geom_smooth(aes(x=DATE,y=WTMP,color="Water"),method="lm", formula="y~x")+
  scale_color_manual("",breaks=c("Air","Water"),values=c("black","blue"))+
  ggtitle("Temperature vs. Time (Monthly)")+
  xlab("Time")+
  ylab("Tempurature")+
  theme(axis.text.x = element_blank())+
  facet_wrap(~factor(MM))
plotFit
```

The coefficients of the linear regression models show the variation of monthly average air temperature and water temperature from 1987 to 2016. 
In January and February, both the air temperature and water temperature decrease. 
In March, April, May, June, August, September, October and December, both the air temperature and water temperature increase. 
In July, the air temperature increases while the water temperature decreases. 
In November, the air temperature decreases while the water temperature increases. 
In most cases, the air temperature and the water temperature is increasing from 1987 to 2016. 
According to National Geographic's artical *Why cold weather doesn't mean climate change is fake*, colder winter could be a signal of extreme weather. 

### Average annual temperature change
We found that temperature decreases in some months. In this step, we are going to find out if it would make a difference to the annual average temperature. Similar as the last step, we are going to plot an average annual temperature versus year graph for visualization. 
```{r}
avg <- MRsub%>%group_by(YYYY)%>%summarize(meanATMP = mean(ATMP), meanWTMP = mean(WTMP))
plotAnnualAvg <- ggplot(data=avg)+
  geom_point(aes(x=YYYY, y=meanATMP), alpha=0.3, color="black")+
  geom_point(aes(x=YYYY, y=meanWTMP), alpha=0.3, color="#56B4E9")+
  geom_smooth(aes(x=YYYY,y=meanATMP,color="Air"),method="lm", formula="y~x")+
  geom_smooth(aes(x=YYYY,y=meanWTMP,color="Water"),method="lm", formula="y~x")+
  scale_color_manual("",breaks=c("Air","Water"),values=c("black","blue"))+
  ggtitle("Average Annual Temperature vs. Year")+
  xlab("Year")+
  ylab("Annual Average Tempurature")
plotAnnualAvg
fitA_annual <- lm(meanATMP~YYYY, data=avg)
fitW_annual <- lm(meanWTMP~YYYY, data=avg)
cat(paste0("By average, the annual average air temperature increases ",
          as.character(coef(fitA_annual)[2]), 
          " celcius degrees \nand the annual average water temperature increases ",
          as.character(coef(fitW_annual)[2]), 
          " celcius degrees every year. ", 
          sep=""))
```

The graph and the coefficient of the linear regression line show that even though temperature decreases in some months, both the air temperature and water temperature is getting warmer in general. The annual average temperatures in 1987, 2010, 2015 and 2016 are also noticeable because of the large difference comparing to the others, which indicates the occurrence of extreme climate. The abrupt change from 2015 to 2016 further demonstrates the unusual climate change. 

### Conclusion
There are evidence validating the climate change in Boston area. Even though the overall temperature is changing slowly, it is gradually increasing in the main direction. The decreasing temperature in winter also prove the extreme climate change. However, the conclusion is only based on temperature change. To further improve the credibility of global warming, we need to increase our sample size and number of predictors. For example, to prove that the climate change is really global, we need to consider about other places in the world. Moreover, temperature change is not the only phenomenon for global warming. Data such as sea levels and amount of carbon dioxide in the ocean could also help us to make a comprehensive analysis on evidence of global warming. 

### Reference
NASA. (2018, September 21). *Climate change evidence: How do we know?* Climate Change: Vital Signs of the Planet. https://climate.nasa.gov/evidence/

Gibbens, S. (2019, January 23). *Why cold weather doesn’t mean climate change is fake*. National Geographic. https://www.nationalgeographic.com/environment/2019/01/climate-change-colder-winters-global-warming-polar-vortex/


